---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## The libraries that will be of support are loaded

```{r, eval=TRUE}
library(ggplot2)
library(dplyr)
library(lubridate)
options(dplyr.summarise.inform = FALSE)
###### Get rid of info of dplyr when grouping: `summarise()` regrouping output by 'species' (override with `.groups` argument)
###### https://rstats-tips.net/2020/07/31/get-rid-of-info-of-dplyr-when-grouping-summarise-regrouping-output-by-species-override-with-groups-argument/

######   2020/05/06
######  dplyr, dplyr-1-0-0
###### Hadley Wickham, Kirill Müller

###### https://www.tidyverse.org/blog/2020/05/dplyr-1-0-0-last-minute-additions/
```


## Loading and preprocessing the initialData

* 1. Load the initialData (i.e. read.csv())
* 2. Process/transform the initialData (if necessary) into a format suitable for your analysis

```{r, echo=TRUE}
unzip("activity.zip")
initialData <- read.csv("activity.csv", colClasses=c("numeric", "Date", "numeric"))
```




## What is mean total number of steps taken per day?

```{r, echo=TRUE}
total_steps <- initialData %>%
  group_by(date) %>%
  summarise(daily_steps = sum(steps, na.rm = TRUE))

ggplot(total_steps, aes(daily_steps)) + 
  geom_histogram(binwidth = 2500, col="darkgreen", fill="lightgreen") +
  ggtitle("Histogram of steps per day") +
  xlab("Total number of steps taken each day") + 
  ylab("Frequency") +
  theme( plot.title = element_text(hjust = 0.5))
```

```{r, echo=TRUE}
mean = mean(total_steps$daily_steps, na.rm=TRUE)
```
```{r, echo=FALSE}
mean
```

```{r, echo=TRUE}
median = median(total_steps$daily_steps, na.rm=TRUE)
```
```{r, echo=FALSE}
median
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
interval_steps <- initialData %>% 
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm =TRUE))

ggplot(data=interval_steps, aes(x=interval, y=steps)) +
    geom_line(col="red") +
    ggtitle("Average steps per time interval") +
    xlab("5-minute intervals") +
    ylab("Average number of steps taken") +
    theme( plot.title = element_text(hjust = 0.5))
```   


## Imputing missing values

##### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, eval=TRUE}
sum(is.na(initialData$steps))
```



```{r, echo=TRUE}
imputed_initialData <- initialData %>%
  mutate(
    steps = case_when(
      is.na(steps) ~ interval_steps$steps[match(initialData$interval, interval_steps$interval)],      
      TRUE ~ as.numeric(steps)
    ))


imputed_total_steps <- imputed_initialData %>% group_by(date) %>% summarise(daily_steps = sum(steps))

ggplot(imputed_total_steps, aes(daily_steps)) + 
  geom_histogram(binwidth = 2500, col="darkgreen", fill="lightgreen") + 
  ggtitle("Total number of steps taken each day") + 
  xlab("Total steps per day") +
  ylab("Frequency") +
  theme( plot.title = element_text(hjust = 0.5))
```

##### We now compute the mean and median number of daily steps of the computed initial Data.


```{r, echo=TRUE}
imputed_mean = mean(imputed_total_steps$daily_steps, na.rm=TRUE)
```
```{r, echo=FALSE}
imputed_mean
```
```{r, echo=TRUE}
imputed_median = median(imputed_total_steps$daily_steps, na.rm=TRUE)
```
```{r, echo=FALSE}
imputed_median
```
##### We can calculate the difference of the means and medians between imputed and original initial Data.


```{r, echo=TRUE}
mean_diff <- imputed_mean - mean 
```
```{r, echo=FALSE}
mean_diff
```
```{r, echo=TRUE}
median_diff <- imputed_median - median
```
```{r, echo=FALSE}
median_diff
```

##### Are there differences in activity patterns between weekdays and weekends?

```{r, eval=TRUE}
library(lubridate)
```

```{r, echo=TRUE}
day_of_week <- imputed_initialData %>%
  mutate(
    date = ymd(date),
    weekday_or_weekend = case_when(wday(date) %in% 2:6 ~ "Weekday",
                                   wday(date) %in% c(1,7) ~ "Weekend")
  ) %>% select(-date) %>%
  group_by(interval, weekday_or_weekend) %>%
  summarise(
    steps = mean(steps), .groups = NULL
  )
###### summarise() and grouping 
###### There's a common confusion about the result of summarise(). How do you think the result of the following code will be grouped?
###### https://www.tidyverse.org/blog/2020/05/dplyr-1-0-0-last-minute-additions/

ggplot(day_of_week, aes(interval, steps)) + 
  geom_line(col="red") +
  facet_wrap(~weekday_or_weekend, nrow = 2) +
  ggtitle("Average daily steps by type of date") + 
  xlab("5-Minute intervals") + 
  ylab("Average number of steps") +
  theme( plot.title = element_text(hjust = 0.5))
```





